---
title: "dplyr"
author: "ricardondata"
date: "2024-10-25"
output: html_document
   coding_folding: "show"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message= FALSE)
```

```{r Load Labraries}
#install.package(tidyverse)
#install.package(sjlabelle)
#install.package(sdcTable)
#install.package(sdcHierarchies)
library(tidyverse) #This will include the "dplyr" package
library(dplyr)
library(tidyr)
library(sjlabelled)
library(sdcTable)
library(sdcHierarchies)
```


```{r load labrary fo ghstat}
remotes::install_github("https://github.com/agyei-boadisly/gssStatsBank")
```


```{r load data}
 View(phc_data)
```


# Selecting the the specific variable for the measuring access to housing using the following indicators #
```{r Accessing the housing }

access_housing<-phc_data %>% 
  select(nqid, pid, region, urbrur, totpopn, restype, weight, a11c, a11d, h01, h08a, h08b, h05, h07a, h06, h10, p02, p12b, p16, p17)

access_housing
  
```
### Note: just the characteristics of the Head "a11c" of Household ###

### tenureship (h05) and Sex (a11d) ###  

# "select()" the type of variable  we want to use 
# "mutate()" the change the variable types into labels
# note: just the categorical variables (dummy), so that we can see the tyoe name 
```{r }
tenure_ship <- access_housing %>% 
  select(nqid, pid, region, urbrur, weight, a11d, a11c, totpopn, restype, h05) %>% 
  mutate(across(c(region, urbrur, a11c, a11d, h05, restype), sjlabelled::as_label))
tenure_ship

```



```{r Selecting the the Head of the household}

tenure_ship_0 <- tenure_ship %>% 
  filter(a11c == "Head") %>% 
  mutate(ownership_type = case_when(h05 %in% c("Owner occupied", "Renting", "Rent-free") ~ "Secure tenancy", TRUE ~ "Not secure tenancy"),
         loc2 = factor(urbrur, levels = c("Urban", "Rural")),
         sex = factor(a11d, levels = c("Male", "Female"))) %>% 
  mutate(ownership_type = factor(ownership_type, levels = c("Secure tenancy", "Not secure tenancy")))
tenure_ship_0
```


```{r ghstat package}
aa <- ghstats::process_data(tenure_ship_0, 
                      var_names = c("ownership_type", "urbrur", "a11d"),
                      var_totals = c("Total", "All localities", "Both sexes"),
                      var_display_names = c("Tenancy agreement", "Locality type", "Sex"),
                      weight_var = "weight",
                      summary_type = "percent"
)

aa

```


```{r}

tenure_ship_0 %>% 
  group_by(ownership_type, a11d, urbrur) 

```




### Ownership of housing (h06) and Sex(a11d)  ###

```{r}
ownership_of_house <- access_housing %>% 
  select(nqid, pid, region, urbrur, weight, a11d, a11c, totpopn, restype, h06) %>% 
  mutate(across(c(region, urbrur, a11c, a11d, h06, restype), sjlabelled::as_label))
ownership_of_house
```


```{r}
ownership_of_house_1 <- ownership_of_house %>% 
  filter(a11c == "Head") %>% 
  mutate(
    specific_owner_house = case_when(
      h06 %in% c("Family property", "Relative not household member") ~ "Family/Relative",
      h06 %in% c("Other private individual", "Other private agency", "Private employer", "Estate developer") ~ "Private",
      h06 %in% c("Public/Government", "Other") ~ "Public/Other",
      is.na(h06) ~ "NA",
      TRUE ~ "Unknown"
      ),
    loc2 = factor(urbrur, levels = c("Urban", "Rural")),
    sex = factor(a11d, levels = c("Male", "Female"))
    ) %>% 
  mutate(specific_owner_house = factor(specific_owner_house, levels = c("Family/Relative", "Private", "Public/Other", "NA", "Unknown"))
  )

ownership_of_house_1
```

```{r ghstat package}
owners <- ghstats::process_data(ownership_of_house_1, 
                      var_names = c("specific_owner_house", "urbrur", "a11d"),
                      var_totals = c("Total", "All localities", "Both sexes"),
                      var_display_names = c("Owns the Place", "Locality type", "Sex"),
                      weight_var = "weight",
                      summary_type = "percent"
)

owners
```



```{r}
owners_1<- ownership_of_house_1 %>% 
  group_by(specific_owner_house, a11d, urbrur) 
owners_1
```

















































